---
title: "Quantitative social science with R"
subtitle: "Get and manipulate data"
author: "Edu Gonzalo Almorox"
date: "18/10/2017"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

exclude: true

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
```

---

# Outline

.pull-left[1. Import data

1. Types of datasets

1. Manipulate data

 
]

.pull-right[

<figure><img src="https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcTW-RXsNKw8mOlCLUu54hp1hMyo6dPIlXD4XiYiAjxBUSamo0y_"><figcaption>.  <a href=""></a></figcaption></figure>

]
---
class: center, middle

# Import data into R
---
## Getting data 

-  There are two possible ways to input information: 

     - **Manually**
     
     - **Import from somewhere**

- The majority of the analyses consist of data created externally 

    - Data are delivered in different formats
    
    - Important to understand how information is structured 
    
    
---
## Step 1: Get started

The first commands consist of the following instructions intended to locate the project in the computer and load the set of functions needed for the analysis. 

* Working directory
```{r, echo = TRUE, eval = FALSE}

# working directory
setwd("~/Users/my-file/my_document") 
```

* Packages
```{r, echo = TRUE, eval = FALSE}
install.packages() # for installing packages 
```

* Library
```{r, echo = TRUE, eval = FALSE}
library() # for loading libraries
```


---
## Step 2: Load your data 

* Various packages and libraries that allow for this. 

* Use depends on the type of data to be loaded

| Function     | Package    | Data 
|---------------|-----------| ------------------------
| `read.csv()`  | `base`    | Comma separated values
| `read.dta()`  | [`foreign`](https://cran.r-project.org/web/packages/foreign/foreign.pdf) | Stata files (before Stata 13)
| `import()`    |  [`rio`](https://cran.r-project.org/web/packages/rio/rio.pdf)   | All types
| `read_csv()`  |  [`readr`](https://cran.r-project.org/web/packages/readr/readr.pdf)  | Comma separated values
| `read_xlsx()` |  [`readr`](https://cran.r-project.org/web/packages/readr/readr.pdf)  | Excel type
| `read_excel()`| [`readxl`](https://cran.r-project.org/web/packages/readxl/readxl.pdf)  | Excel files
| `read_xml()`  | [`xml2`](https://cran.r-project.org/web/packages/xml2/xml2.pdf)        | XML files


---
## Load data: read.csv




---
class: center, middle

# Thanks!

@EduGonzalo

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

